---
title: "raw_data"
author: "Jueun Park"
date: "2024-11-07"
output: html_document
---

```{r}
library(ggplot2)
library(plyr) # join all
library(dplyr)
library(tidyr)
library(openxlsx)
library(readxl)
library(stringr) #string separate 
library(ggthemes)
library(stringr)
library(broom)
library(glmnet)
library(plm) #panel data regression
library(car) #VIF
library(forecast) #prediction
library(lubridate) #date
library(purrr)
library(mice)
library(zoo)
```


#1. Hate Crime Report data

```{r}
#State abbreviation data

state_abbrev <- read.csv("data_raw/state_abbrev.csv")
state_abbrev <- state_abbrev %>%
  dplyr::rename(state = State,
                state_abbrev = Abbreviation)
```

```{r setup, include=FALSE}
#Cleaning hate crime report data

#Read hate crime report data
hate_raw <- read.csv("data_raw/hate_crime.csv")

#Rename columns and add column of state abbreviation
hate_master <- hate_raw %>%
  dplyr::rename(year = data_year,state = state_name) %>%
  left_join(state_abbrev, by=("state"))


# Categorize sub bias type into bigger category and edit incident date column
hate_master <- hate_master %>%
  mutate(
    bias_desc = str_to_lower(bias_desc),
    bias_category = case_when(
      str_detect(bias_desc, "black|white|asian|hispanic|arab|american indian") ~ "Race/Ethnicity",
      str_detect(bias_desc, "jewish|muslim|catholic|protestant|hindu|sikh") ~ "Religion",
      str_detect(bias_desc, "gay|lesbian|transgender|lgbtq|bisexual") ~ "Sexual Orientation/Gender Identity",
      str_detect(bias_desc, "physical disability|mental disability") ~ "Disability",
      str_detect(bias_desc, "heterosexual|gender non-conforming|male|female") ~ "Other Social Groups",
      TRUE ~ "Intersectional Bias"
    ),
    incident_date = as.Date(incident_date, format = "%Y-%m-%d"),
         year = year(incident_date),
         month = month(incident_date, label = TRUE)
  ) %>%
  select(incident_id, state, state_abbr, region_name, year, month, offender_race, offender_ethnicity, victim_count, total_individual_victims, offense_name, bias_desc, bias_desc, bias_category)

# Save hate master data
write.csv(hate_master, file = "data_created/hate_master.csv", sep = "\t", row.names = F)
```

#2. GDP
```{r}
#Cleaning GDP data

#Read gdp dataset
gdp_raw <- read.csv("data_raw/GDP_state.csv", skip=3)

#Rename columns
gdp_master <- dplyr::rename(gdp_raw,
                   state = GeoName,
                   "2016" = X2016,
                   "2017" = X2017,
                   "2018" = X2018,
                   "2019" = X2019,
                  "2020" = X2020,
                  "2021" = X2021,
                  "2022" = X2022,
                  "2023" = X2023)

#Remove column that is not needed
gdp_master = subset(gdp_master, select = -c(GeoFips)) 

#Final Tidy dataset for GDP
gdp_master <- pivot_longer(
  gdp_master,
  cols = -state, 
  names_to = "year", 
  values_to = "gdp"
)

# Save gdp cleaned master data
write.csv(gdp_master, file = "data_created/gdp_master.csv", sep = "\t", row.names = F)
```


#3.Political dataset
```{r}
#Cleaning Political data

political_raw <- read.xlsx("data_raw/political_state.xlsx", sheet = "16. Appendix 1A")

#Election result - 2020
result_2020 <- political_raw %>% select("Appendix.A:.1996-2020.Presidential.General.Election", X2, X3)
result_2020 <- result_2020[-(1:5),]

result_2020 <- dplyr::rename(result_2020,
                      state = "Appendix.A:.1996-2020.Presidential.General.Election",
                      Democratic = X2,
                      Republican = X3)

result_2020 <- result_2020 %>%
  mutate(
    Democratic = as.numeric(Democratic),
    Republican = as.numeric(Republican),
    "2020" = ifelse(Democratic > Republican, "Democratic", "Republican")
  )


#Election result - 2016
result_2016 <- political_raw %>% select("Appendix.A:.1996-2020.Presidential.General.Election", X4, X5)
result_2016 <- result_2016[-(1:5),]

result_2016 <- dplyr::rename(result_2016,
                      state = "Appendix.A:.1996-2020.Presidential.General.Election",
                      Democratic = X4,
                      Republican = X5) %>%
  mutate(
    Democratic = as.numeric(Democratic),
    Republican = as.numeric(Republican),
    "2016" = ifelse(Democratic > Republican, "Democratic", "Republican")
  )

# Total Election result 2016-2020
# Final Tidy dataset for political leaning
political_master <- left_join(result_2020, result_2016, by=("state")) %>%
  select(state, "2020", "2016") %>%
  pivot_longer(
  cols = -state, 
  names_to = "year", 
  values_to = "political"
) %>%
  filter(row_number() <= n()-6)

# Save political cleaned master data
write.csv(political_master, file = "data_created/political_master.csv", sep = "\t", row.names = F)
```


#4.Population Density
```{r}
#Cleaning Population density data


# Read files for this variable
land_area <- read.csv("data_raw/state_landArea.csv")
state_pop <- read_excel("data_raw/population_state.xls", sheet = "Annual")



#Extract only year
state_pop$DATE <- gsub("-01-01","", as.character(state_pop$DATE))

#Tidy data
popDens_master<- pivot_longer(
  state_pop,
  cols = -DATE, 
  names_to = "state_abbrev", 
  values_to = "population"
)

#Extract only state abbreviation
popDens_master$state_abbrev <- gsub("POP","", as.character(popDens_master$state_abbrev))


#Join necessary dataset
popDens_master <- left_join(popDens_master, state_abbrev, by=("state_abbrev"))
popDens_master <- left_join(popDens_master, land_area, by=("state"))

#Final dataset for population density
popDens_master <- popDens_master %>%
  mutate(density = round((population*1000)/LandArea), digits = 2) %>%
  select(state, state_abbrev, DATE, density) %>%
  dplyr::rename(year = DATE)

# Save cleaned master data for population density
write.csv(popDens_master, file = "data_created/popDens_master.csv", sep = "\t", row.names = F)
```

#5. Unemployment Rate
```{r}
#Cleaning Unemployment data

#read uemployment data by state
unemp_raw <- read_xlsx("data_raw/unemployment_state.xlsx", skip = 6)

#Final Tidy unemployment data
unemp_master <- unemp_raw %>%
  select("...2","Labor force...3", "rate") %>%
  dplyr::rename(state ="...2",
         year = "Labor force...3",
         unemp_rate = "rate") %>%
  filter(year > 2015) %>%
  group_by(state, year) %>% # calculate the mean unemployment rate for the year
  dplyr::summarize(avg_unemp = (round(mean(unemp_rate),2))/100, .groups = "keep")

# Save cleaned master data for unemployment rate
write.csv(unemp_master, file = "data_created/unemp_master.csv", sep = "\t", row.names = F)
```

#6.Gender distribution
```{r}
#Cleaning gender distribution data

#Reading all the dataset for gender distribution(2016-2023)
gender_2023 <- read.csv("data_raw/sex_2023.csv", skip = 2) %>%
  select(Location, Male, Female) %>%
  dplyr::rename(state = Location,
         "Male 2023" = Male,
         "Female 2023" = Female)

gender_2022 <- read.csv("data_raw/sex_2022.csv", skip = 2) %>%
  select(Location, Male, Female) %>%
  dplyr::rename(state = Location,
         "Male 2022" = Male,
         "Female 2022" = Female)

gender_2021 <- read.csv("data_raw/sex_2021.csv", skip = 2) %>%
  select(Location, Male, Female) %>%
  dplyr::rename(state = Location,
         "Male 2021" = Male,
         "Female 2021" = Female)

gender_2019 <- read.csv("data_raw/sex_2019.csv", skip = 2) %>%
  select(Location, Male, Female) %>%
  dplyr::rename(state = Location,
         "Male 2019" = Male,
         "Female 2019" = Female)

gender_2018 <- read.csv("data_raw/sex_2018.csv", skip = 2) %>%
  select(Location, Male, Female) %>%
  dplyr::rename(state = Location,
         "Male 2018" = Male,
         "Female 2018" = Female)

gender_2017 <- read.csv("data_raw/sex_2017.csv", skip = 2) %>%
  select(Location, Male, Female) %>%
  dplyr::rename(state = Location,
         "Male 2017" = Male,
         "Female 2017" = Female)

gender_2016 <- read.csv("data_raw/sex_2016.csv", skip = 2) %>%
  select(Location, Male, Female) %>%
  dplyr::rename(state = Location,
         "Male 2016" = Male,
         "Female 2016" = Female)


#Join all the dataset
gender_raw <- join_all(list(gender_2016, gender_2017, gender_2018, gender_2019, gender_2021, gender_2022, gender_2023), by=("state"), type = "left")


#Final tidy gender distribution data
gender_master <- pivot_longer(
  gender_raw,
  cols = -state, 
  names_to = "gender", 
  values_to = "gender_dist"
)

gender_master[c('gender', 'year')] <- str_split_fixed(gender_master$gender, ' ', 2)

# Save cleaned master data for gender distribution
write.csv(gender_master, file = "data_created/gender_master.csv", sep = "\t", row.names = F)
```

#7. Race Distribtuion
```{r}
#Cleaning race distribution data

#Reading all the dataset for race distribution(2016-2023)
race_2023 <- read.csv("data_raw/race_2023.csv", skip = 2) %>%
  select(Location, White, Black, Hispanic, Asian) %>%
  dplyr::rename(state = Location,
         "White 2023" = White,
         "Black 2023" = Black,
         "Hispanic 2023" = Hispanic,
         "Asian 2023" = Asian)

race_2022 <- read.csv("data_raw/race_2022.csv", skip = 2) %>%
  select(Location, White, Black, Hispanic, Asian) %>%
  dplyr::rename(state = Location,
         "White 2022" = White,
         "Black 2022" = Black,
         "Hispanic 2022" = Hispanic,
         "Asian 2022" = Asian)

race_2021 <- read.csv("data_raw/race_2021.csv", skip = 2) %>%
  select(Location, White, Black, Hispanic, Asian) %>%
  dplyr::rename(state = Location,
         "White 2021" = White,
         "Black 2021" = Black,
         "Hispanic 2021" = Hispanic,
         "Asian 2021" = Asian)

race_2019 <- read.csv("data_raw/race_2019.csv", skip = 2) %>%
  select(Location, White, Black, Hispanic, Asian) %>%
  dplyr::rename(state = Location,
         "White 2019" = White,
         "Black 2019" = Black,
         "Hispanic 2019" = Hispanic,
         "Asian 2019" = Asian)

race_2018 <- read.csv("data_raw/race_2018.csv", skip = 2) %>%
  select(Location, White, Black, Hispanic, Asian) %>%
  dplyr::rename(state = Location,
         "White 2018" = White,
         "Black 2018" = Black,
         "Hispanic 2018" = Hispanic,
         "Asian 2018" = Asian)

race_2017 <- read.csv("data_raw/race_2017.csv", skip = 2) %>%
  select(Location, White, Black, Hispanic, Asian) %>%
  dplyr::rename(state = Location,
         "White 2017" = White,
         "Black 2017" = Black,
         "Hispanic 2017" = Hispanic,
         "Asian 2017" = Asian)

race_2016 <- read.csv("data_raw/race_2016.csv", skip = 2) %>%
  select(Location, White, Black, Hispanic, Asian) %>%
  dplyr::rename(state = Location,
         "White 2016" = White,
         "Black 2016" = Black,
         "Hispanic 2016" = Hispanic,
         "Asian 2016" = Asian)
  
race_2016$"Black 2016" <-gsub("<","0.",as.numeric(race_2016$"Black 2016"))
race_2017$"Black 2017" <-gsub("<","0.",as.numeric(race_2017$"Black 2017"))
race_2018$"Black 2018" <-gsub("<","0.",as.numeric(race_2018$"Black 2018"))


#Join all the dataset
race_master <- join_all(list(race_2016, race_2017, race_2018, race_2019, race_2021, race_2022, race_2023), by=("state"), type = "left")

# Clean data
race_master <- race_master[-(54:65),]
race_master[race_master == "N/A"] <- NA

race_master$"Black 2016" <- as.numeric(as.character(race_master$"Black 2016"))
race_master$"Asian 2016" <- as.numeric(as.character(race_master$"Asian 2016"))
race_master$"Black 2017" <- as.numeric(as.character(race_master$"Black 2017"))
race_master$"Asian 2017" <- as.numeric(as.character(race_master$"Asian 2017"))
race_master$"Black 2018" <- as.numeric(as.character(race_master$"Black 2018"))
race_master$"Asian 2018" <- as.numeric(as.character(race_master$"Black 2018"))
race_master$"Black 2019" <- as.numeric(as.character(race_master$"Black 2019"))
race_master$"Asian 2019" <- as.numeric(as.character(race_master$"Black 2019"))
race_master$"Asian 2021" <- as.numeric(as.character(race_master$"Black 2021"))
race_master$"Asian 2023" <- as.numeric(as.character(race_master$"Black 2023"))

#Final tidy gender distribution data
race_master <- race_master %>%
  pivot_longer(
  cols = -state, 
  names_to = "race", 
  values_to = "race_dist"
)

race_master[c('race', 'year')] <- str_split_fixed(race_master$race, ' ', 2)

# Save cleaned master data for race density
write.csv(race_master, file = "data_created/race_master.csv", sep = "\t", row.names = F)
```


#8. Age distribution
```{r}
#Cleaning age distribution data

#Reading all the dataset for age distribution(2016-2023)
age_2023 <- read.csv("data_raw/age_2023.csv", skip = 2) %>%
  select(Location, Children.0.18, Adults.19.25, Adults.26.34,Adults.35.54, Adults.55.64,X65.) %>%
  dplyr::rename(state = Location,
               "0-18 2023" = Children.0.18,
               "19-25 2023" = Adults.19.25,
               "26-34 2023" = Adults.26.34,
                "35-54 2023" = Adults.35.54,
                "55-64 2023" = Adults.55.64,
                "65+ 2023" = X65.)

age_2022 <- read.csv("data_raw/age_2022.csv", skip = 2) %>%
  select(Location, Children.0.18, Adults.19.25, Adults.26.34,Adults.35.54, Adults.55.64,X65.) %>%
  dplyr::rename(state = Location,
               "0-18 2022" = Children.0.18,
               "19-25 2022" = Adults.19.25,
               "26-34 2022" = Adults.26.34,
                "35-54 2022" = Adults.35.54,
                "55-64 2022" = Adults.55.64,
                "65+ 2022" = X65.)

age_2021 <- read.csv("data_raw/age_2021.csv", skip = 2) %>%
  select(Location, Children.0.18, Adults.19.25, Adults.26.34,Adults.35.54, Adults.55.64,X65.) %>%
  dplyr::rename(state = Location,
               "0-18 2021" = Children.0.18,
               "19-25 2021" = Adults.19.25,
               "26-34 2021" = Adults.26.34,
                "35-54 2021" = Adults.35.54,
                "55-64 2021" = Adults.55.64,
                "65+ 2021" = X65.)

age_2019 <- read.csv("data_raw/age_2019.csv", skip = 2) %>%
  select(Location, Children.0.18, Adults.19.25, Adults.26.34,Adults.35.54, Adults.55.64,X65.) %>%
  dplyr::rename(state = Location,
               "0-18 2019" = Children.0.18,
               "19-25 2019" = Adults.19.25,
               "26-34 2019" = Adults.26.34,
                "35-54 2019" = Adults.35.54,
                "55-64 2019" = Adults.55.64,
                "65+ 2019" = X65.)

age_2018 <- read.csv("data_raw/age_2018.csv", skip = 2) %>%
  select(Location, Children.0.18, Adults.19.25, Adults.26.34,Adults.35.54, Adults.55.64,X65.) %>%
  dplyr::rename(state = Location,
               "0-18 2018" = Children.0.18,
               "19-25 2018" = Adults.19.25,
               "26-34 2018" = Adults.26.34,
                "35-54 2018" = Adults.35.54,
                "55-64 2018" = Adults.55.64,
                "65+ 2018" = X65.)

age_2017 <- read.csv("data_raw/age_2017.csv", skip = 2) %>%
  select(Location, Children.0.18, Adults.19.25, Adults.26.34,Adults.35.54, Adults.55.64,X65.) %>%
  dplyr::rename(state = Location,
               "0-18 2017" = Children.0.18,
               "19-25 2017" = Adults.19.25,
               "26-34 2017" = Adults.26.34,
                "35-54 2017" = Adults.35.54,
                "55-64 2017" = Adults.55.64,
                "65+ 2017" = X65.)

age_2016 <- read.csv("data_raw/age_2016.csv", skip = 2) %>%
  select(Location, Children.0.18, Adults.19.25, Adults.26.34,Adults.35.54, Adults.55.64,X65.) %>%
  dplyr::rename(state = Location,
               "0-18 2016" = Children.0.18,
               "19-25 2016" = Adults.19.25,
               "26-34 2016" = Adults.26.34,
                "35-54 2016" = Adults.35.54,
                "55-64 2016" = Adults.55.64,
                "65+ 2016" = X65.)

#Join all the dataset
age_raw <- join_all(list(age_2016, age_2017, age_2018, age_2019, age_2021, age_2022, age_2023), by=("state"), type = "left")    


#Final tidy age distribution data
age_master <- age_raw %>%
  pivot_longer(
  cols = -state, 
  names_to = "age", 
  values_to = "age_dist"
)


age_master[c('age', 'year')] <- str_split_fixed(age_master$age, ' ', 2)   


# Save cleaned master data for age distribution
write.csv(age_master, file = "data_created/age_master.csv", sep = "\t", row.names = F)
```
#9. Poverty rate
```{r}
# poverty rate

#Read poverty data
poverty_2023 <- read.csv("data_raw/poverty_2023.csv", skip = 2) %>%
  select(state = Location, poverty = Total) %>%
  mutate(year = 2023) %>%
  slice(1:53)


poverty_2022 <- read.csv("data_raw/poverty_2022.csv", skip = 2) %>%
  select(state = Location, poverty = Total) %>%
  mutate(year = 2022) %>%
  slice(1:53)
  
poverty_2021 <- read.csv("data_raw/poverty_2021.csv", skip = 2) %>%
  select(state = Location, poverty = Total) %>%
  mutate(year = 2021) %>%
  slice(1:53)

poverty_2019 <- read.csv("data_raw/poverty_2019.csv", skip = 2) %>%
  select(state = Location, poverty = Total) %>%
  mutate(year = 2019) %>%
  slice(1:53)

poverty_2018 <- read.csv("data_raw/poverty_2018.csv", skip = 2) %>%
  select(state = Location, poverty = Total) %>%
  mutate(year = 2018) %>%
  slice(1:53)

poverty_2017 <- read.csv("data_raw/poverty_2017.csv", skip = 2) %>%
  select(state = Location, poverty = Total) %>%
  mutate(year = 2017) %>%
  slice(1:53)


poverty_2016 <- read.csv("data_raw/poverty_2016.csv", skip = 2) %>%
  select(state = Location, poverty = Total) %>%
  mutate(year = 2016) %>%
  slice(1:53)


#Join all the dataset
poverty_master <- bind_rows(poverty_2016, poverty_2017, poverty_2018, poverty_2019, poverty_2021, poverty_2022, poverty_2023)

# Save cleaned master data for poverty rate
write.csv(poverty_master, file = "data_created/poverty_master.csv", sep = "\t", row.names = F)
```

#10.Education level
```{r}
#education

education_raw <- read_excel("data_raw/education_state.xls", sheet = "Annual")

#Extract only year
education_raw$DATE <- gsub("-01-01","", as.character(education_raw$DATE))

#Tidy data
education_master <- pivot_longer(
  education_raw,
  cols = -DATE, 
  names_to = "state_abbrev", 
  values_to = "college_degree_rate"
)

#Extract only state abbreviation
education_master$state_abbrev <- gsub("GCT1502","", as.character(education_master$state_abbrev))

education_master <- left_join(education_master, state_abbrev, by=("state_abbrev")) %>%
  select(state, year = DATE, college_degree_rate) %>%
  mutate(college_degree_rate = college_degree_rate/100)

# Save cleaned master data for Bachelor degree rate
write.csv(education_master, file = "data_created/education_master.csv", sep = "\t", row.names = F)
```

#11. Covid confirmed cases
```{r}
#COVID cases

covid_master <- read.csv("data_raw/covid_state.csv")  %>%
  mutate(date_updated = as.Date(date_updated, format = "%m/%d/%Y"),
         year = year(date_updated),
         month = month(date_updated, label = TRUE)) %>%
  select(state, year, month, new_cases, tot_cases) %>%
  group_by(state, year, month) %>%
  dplyr::summarise(cum_total_cases = max(tot_cases, na.rm = TRUE),
            monthly_new_cases = sum(new_cases, na.rm = TRUE),
            .groups = "keep") %>%
  dplyr::rename(state_abbrev = state) %>%
  left_join(state_abbrev, by=("state_abbrev"))

# Save cleaned master data for Covid confirmed cases
write.csv(covid_master, file = "data_created/covid_master.csv", sep = "\t", row.names = F)
```

#12. Health status
```{r}
#health


health_2016 <- read.csv("data_raw/health_2016.csv", skip = 2) %>%
  select(state = Location, health_status = All.Adults) %>%
  mutate(year = 2016) %>%
  slice(1:55)

health_2017 <- read.csv("data_raw/health_2017.csv", skip = 2) %>%
  select(state = Location, health_status = All.Adults) %>%
  mutate(year = 2017,
         health_status = as.numeric(health_status)) %>%
  slice(1:55)

health_2018 <- read.csv("data_raw/health_2018.csv", skip = 2) %>%
  select(state = Location, health_status = All.Adults) %>%
  mutate(year = 2018,
         health_status = as.numeric(health_status)) %>%
  slice(1:55)

health_2019 <- read.csv("data_raw/health_2019.csv", skip = 2) %>%
  select(state = Location, health_status = All.Adults) %>%
  mutate(year = 2019,
         health_status = as.numeric(health_status)) %>%
  slice(1:55)

health_2020 <- read.csv("data_raw/health_2020.csv", skip = 2) %>%
  select(state = Location, health_status = All.Adults) %>%
  mutate(year = 2020,
         health_status = as.numeric(health_status)) %>%
  slice(1:55)

health_2021 <- read.csv("data_raw/health_2021.csv", skip = 2) %>%
  select(state = Location, health_status = All.Adults) %>%
  mutate(year = 2021,
         health_status = as.numeric(health_status)) %>%
  slice(1:55)

health_2022 <- read.csv("data_raw/health_2022.csv", skip = 2) %>%
  select(state = Location, health_status = All.Adults) %>%
  mutate(year = 2022,
         health_status = as.numeric(health_status)) %>%
  slice(1:55)

health_master <- bind_rows(health_2016, health_2017, health_2018, health_2019, health_2020, health_2021, health_2022)

# Save cleaned master data for health status
write.csv(health_master, file = "data_created/health_master.csv", sep = "\t", row.names = F)
```

#Data for Fixed effect model
```{r message=FALSE, warning=FALSE}
# Identify which social factors were most closely associated with the increase in hate crimes

# List of vulnerable states
vul_state <- c("Arizona", "California", "Hawaii", "Idaho", "Indiana", "Louisiana", "Minnesota", "Mississippi", "Nebraska", "Nevada", "New Jersey", "New York",	"Pennsylvania", "Washington")
```


#Hate data
```{r message=FALSE, warning=FALSE}
# Hate crime data
fe_hate <- read.csv("data_created/hate_master.csv") %>%
  filter(year > 2019) %>%
  filter(state %in% vul_state) %>%
  group_by(year, state) %>%
  dplyr::summarize(total_hate_crimes = n(), .groups = "keep")
```


#Age distribution data
```{r message=FALSE, warning=FALSE}
fe_age <- read.csv("data_created/age_master.csv") %>%
  filter(state %in% vul_state) %>%
  complete(state, age, year = 2016:2023, fill = list(age_dist = NA)) %>%
  arrange(state, age, year) %>%
  group_by(state, age) %>%
  mutate(age_dist = na.approx(age_dist, na.rm = FALSE)) %>% 
  ungroup()

fe_age <- fe_age %>%
  filter(year >= 2020) %>%  
  pivot_wider(names_from = age, values_from = age_dist) %>%
  mutate(
    young = `0-18` + `19-25`,
    middle_Aged = `26-34` + `35-54`,
    old = `55-64` + `65+`
  ) %>%
  select(state, year, young, middle_Aged, old)
```


#Race distribution data
```{r message=FALSE, warning=FALSE}
fe_race <- read.csv("data_created/race_master.csv") %>%
  filter(state %in% vul_state) %>%
  complete(state, race, year = 2016:2023, fill = list(race_dist = NA)) %>%
  arrange(state, race, year) %>%
  group_by(state, race) %>%
  mutate(race_dist = na.approx(race_dist, na.rm = FALSE)) %>% 
  ungroup() %>%
  filter(year >= 2020) %>%
  pivot_wider(names_from = race, values_from = race_dist)
```


#Gender
```{r message=FALSE, warning=FALSE}
fe_gender <- read.csv("data_created/gender_master.csv") %>%
  filter(state %in% vul_state) %>%
  complete(state, gender, year = 2016:2023, fill = list(gender_dist = NA)) %>%
  arrange(state, gender, year) %>%
  group_by(state, gender) %>%
  mutate(gender_dist = na.approx(gender_dist, na.rm = FALSE)) %>%
  ungroup() %>%
  filter(year >= 2020) %>%
  pivot_wider(names_from = gender, values_from = gender_dist)
```


#GDP
```{r message=FALSE, warning=FALSE}
fe_gdp <- read.csv("data_created/gdp_master.csv") %>%
  filter(state %in% vul_state) %>%
  filter(year >= 2020)
```


#Density
```{r message=FALSE, warning=FALSE}
fe_density <- read.csv("data_created/popDens_master.csv") %>%
  filter(state %in% vul_state) %>%
  filter(year >= 2020) %>%
  select(-state_abbrev)
```


#Political
```{r message=FALSE, warning=FALSE}
fe_poitical <- read.csv("data_created/political_master.csv") %>%
  complete(state, year = 2016:2023) %>%
  arrange(state, year) %>%
  group_by(state) %>%
  fill(political, .direction = "downup") %>%
  ungroup() %>%
  filter(state %in% vul_state) %>%
  filter(year >= 2020)
```


#Unemployment Rate
```{r message=FALSE, warning=FALSE}
fe_unemployment <- read.csv("data_created/unemp_master.csv") %>%
  filter(year >= 2020 & year < 2024) %>%
  filter(state %in% vul_state)
```


#Health status
```{r message=FALSE, warning=FALSE}
fe_health <- read.csv("data_created/health_master.csv") %>%
  filter(state %in% vul_state) %>%
  complete(state, year = 2016:2023, fill = list(health_status = NA)) %>%
  arrange(state, year) %>%
  group_by(state) %>%
  mutate(health_status = na.approx(health_status, na.rm = FALSE, rule = 2)) %>% 
  ungroup() %>%
  filter(year >= 2020)
```


#Poverty rate
```{r message=FALSE, warning=FALSE}
fe_poverty <- read.csv("data_created/poverty_master.csv") %>%
  filter(state %in% vul_state) %>%
  complete(state, year = 2016:2023, fill = list(poverty = NA)) %>%
  arrange(state, year) %>%
  group_by(state) %>%
  mutate(poverty = na.approx(poverty, na.rm = FALSE)) %>% 
  ungroup() %>%
  filter(year >= 2020)
```


# Education(Bachelor degree attained rate)
```{r message=FALSE, warning=FALSE}
fe_education <- read.csv("data_created/education_master.csv") %>%
  filter(state %in% vul_state) %>%
  filter(year >= 2020)
```

```{r message=FALSE, warning=FALSE}
fe_data <- plyr::join_all(
  list(fe_hate, fe_age, fe_race, fe_gender, fe_gdp, fe_density, fe_poitical, fe_unemployment, fe_health, fe_poverty, fe_education),
  by=c("state","year"),
  type="left")

# Save fixed effect model master data
write.csv(fe_data, file = "data_created/fe_data.csv", sep = "\t", row.names = F)
```

---


#Data for logistic regression


#Hate data
```{r message=FALSE, warning=FALSE}
# Hate crime data
logi_hate <- read.csv("data_created/hate_master.csv") %>%
  filter(year > 2019) %>%
  group_by(year, state) %>%
  dplyr::summarize(total_hate_crimes = n(), .groups = "keep")
```


#Age distribution data
```{r message=FALSE, warning=FALSE}
logi_age <- read.csv("data_created/age_master.csv") %>%
  complete(state, age, year = 2016:2023, fill = list(age_dist = NA)) %>%
  arrange(state, age, year) %>%
  group_by(state, age) %>%
  mutate(age_dist = na.approx(age_dist, na.rm = FALSE)) %>% 
  ungroup()

logi_age <- logi_age %>%
  filter(year >= 2020) %>% 
  pivot_wider(names_from = age, values_from = age_dist) %>%
  mutate(
    young = `0-18` + `19-25`,
    middle_Aged = `26-34` + `35-54`,
    old = `55-64` + `65+`
  ) %>%
  select(state, year, young, middle_Aged, old)
```


#Race distribution data
```{r message=FALSE, warning=FALSE}
logi_race <- read.csv("data_created/race_master.csv") %>%
  complete(state, race, year = 2016:2023, fill = list(race_dist = NA)) %>%
  arrange(state, race, year) %>%
  group_by(state, race) %>%
  mutate(race_dist = na.approx(race_dist, na.rm = FALSE)) %>% 
  ungroup() %>%
  filter(year >= 2020) %>%
  pivot_wider(names_from = race, values_from = race_dist)
```


#Gender
```{r message=FALSE, warning=FALSE}
logi_gender <- read.csv("data_created/gender_master.csv") %>%
  complete(state, gender, year = 2016:2023, fill = list(gender_dist = NA)) %>%
  arrange(state, gender, year) %>%
  group_by(state, gender) %>%
  mutate(gender_dist = na.approx(gender_dist, na.rm = FALSE)) %>% 
  ungroup() %>%
  filter(year >= 2020) %>%
  pivot_wider(names_from = gender, values_from = gender_dist)
```


#GDP
```{r message=FALSE, warning=FALSE}
logi_gdp <- read.csv("data_created/gdp_master.csv") %>%
  filter(year >= 2020)
```


#Density
```{r message=FALSE, warning=FALSE}
logi_density <- read.csv("data_created/popDens_master.csv") %>%
  filter(year >= 2020) %>%
  select(-state_abbrev)
```


#Political
```{r message=FALSE, warning=FALSE}
logi_poitical <- read.csv("data_created/political_master.csv") %>%
  complete(state, year = 2016:2023) %>%
  arrange(state, year) %>%
  group_by(state) %>%
  fill(political, .direction = "downup") %>%
  ungroup() %>%
  filter(year >= 2020)
```


#Unemployment Rate
```{r message=FALSE, warning=FALSE}
logi_unemployment <- read.csv("data_created/unemp_master.csv") %>%
  filter(year >= 2020 & year < 2024)
```


#Health status
```{r message=FALSE, warning=FALSE}
logi_health <- read.csv("data_created/health_master.csv") %>%
  complete(state, year = 2016:2023, fill = list(health_status = NA)) %>%
  arrange(state, year) %>%
  group_by(state) %>%
  mutate(health_status = na.approx(health_status, na.rm = FALSE, rule = 2)) %>% 
  ungroup() %>%
  filter(year >= 2020)
```


#Poverty rate
```{r message=FALSE, warning=FALSE}
logi_poverty <- read.csv("data_created/poverty_master.csv") %>%
  complete(state, year = 2016:2023, fill = list(poverty = NA)) %>%
  arrange(state, year) %>%
  group_by(state) %>%
  mutate(poverty = na.approx(poverty, na.rm = FALSE)) %>% 
  ungroup() %>%
  filter(year >= 2020)
```


# Education(Bachelor degree attained rate)
```{r message=FALSE, warning=FALSE}
logi_education <- read.csv("data_created/education_master.csv") %>%
  filter(year >= 2020)
```

```{r message=FALSE, warning=FALSE}
logi_data <- plyr::join_all(
  list(logi_hate, logi_age, logi_race, logi_gender, logi_gdp, logi_density, logi_poitical, logi_unemployment, logi_health,
       logi_poverty, logi_education),
  by=c("state","year"),
  type="left")

# Save fixed effect model master data
write.csv(logi_data, file = "data_created/logi_data.csv", sep = "\t", row.names = F)
```



